---
title: "Home Work Assignment - 01 - Test/ Evaluation Data"
author: "Critical Thinking Group 5"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---


```{r, echo = TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
library(dplyr)
library(psych)
library(stringr)
library(car)


url <- "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW1/moneyball-evaluation-data.csv"
moneyball<- read.csv(url)
head(moneyball)
moneyball2<- select(moneyball, -(INDEX))
summary(moneyball2)

```


# Data Exploration


## Summary / Descriptives / Correlation

```{r, echo = TRUE, warning=FALSE, message=FALSE}
ds_stats <- psych::describe(moneyball2, skew = FALSE, na.rm = TRUE)[c(3:6)]
ds_stats <- cbind(VARIABLE_NAME = rownames(ds_stats), ds_stats)
#rownames(ds_stats) <- NULL

Variable<- rownames(ds_stats)

fun <- function(x) sum(!complete.cases(x))
Missing <- sapply(moneyball2[Variable], FUN = fun) 


```


## Distribution and Correlation

```{r}
show_charts <- function(x, ...) {
    
    par(mfrow=c(2,3))
    
    xlabel <- unlist(str_split(deparse(substitute(x)), pattern = "\\$"))[2]
#    ylabel <- unlist(str_split(deparse(substitute(y)), pattern = "\\$"))[2]
    
    hist(x,main=xlabel)
    boxplot(x,main=xlabel)

    y<-log(x)
    boxplot(y,main='log transform')
    y<-sqrt(x)
    boxplot(y,main='sqrt transform')
    y<-sin(x)
    boxplot(y,main='sin transform')
    y<-(x)^(1/9)
    boxplot(y,main='ninth transform')
}


show_charts(moneyball2$TEAM_BATTING_H)

show_charts(moneyball2$TEAM_BATTING_2B)

show_charts(moneyball2$TEAM_BATTING_3B)

show_charts(moneyball2$TEAM_BATTING_HR)

show_charts(moneyball2$TEAM_BATTING_BB)

show_charts(moneyball2$TEAM_BATTING_SO)

show_charts(moneyball2$TEAM_BASERUN_SB)

show_charts(moneyball2$TEAM_FIELDING_E)

show_charts(moneyball2$TEAM_FIELDING_DP)

show_charts(moneyball2$TEAM_PITCHING_BB)

show_charts(moneyball2$TEAM_PITCHING_H)

show_charts(moneyball2$TEAM_PITCHING_HR)

show_charts(moneyball2$TEAM_PITCHING_SO)
```


# Data Preparation

Now that we have the preliminary analysis ready, we will go ahead and carry out the necessary transformations to the data. 

This will primarily take care of Missing Values, Handle Outliers and create some additional variables.
\
\

## Outliers 

For outliers, we will create 2 sets of variables. 

The first set uses the capping method. In this method, we will replace all outliers that lie outside the 1.5 times of IQR limits. We will cap it by replacing those observations less than the lower limit with the value of 5th %ile and those that lie above the upper limit with the value of 95th %ile. 

Accordingly we create the following new variables while retaining the original variables as is. 

TEAM_BATTING_H_NEW
TEAM_BATTING_2B_NEW
TEAM_BATTING_3B_NEW
TEAM_BATTING_BB_NEW
TEAM_BASERUN_SB_NEW
TEAM_FIELDING_E_NEW
TEAM_FIELDING_DP_NEW
TEAM_PITCHING_BB_NEW
TEAM_PITCHING_H_NEW
TEAM_PITCHING_HR_NEW
TEAM_PITCHING_SO_NEW


```{r}
# function for removing outliers - http://r-statistics.co/Outlier-Treatment-With-R.html

treat_outliers <- function(x) {
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
 
return(x)
}

moneyball2$TEAM_BATTING_H_NEW <- treat_outliers(moneyball2$TEAM_BATTING_H)
moneyball2$TEAM_BATTING_2B_NEW <- treat_outliers(moneyball2$TEAM_BATTING_2B)
moneyball2$TEAM_BATTING_3B_NEW <- treat_outliers(moneyball2$TEAM_BATTING_3B)
moneyball2$TEAM_BATTING_BB_NEW <- treat_outliers(moneyball2$TEAM_BATTING_BB)
moneyball2$TEAM_BASERUN_SB_NEW <- treat_outliers(moneyball2$TEAM_BASERUN_SB)
moneyball2$TEAM_FIELDING_E_NEW <- treat_outliers(moneyball2$TEAM_FIELDING_E)
moneyball2$TEAM_FIELDING_DP_NEW <- treat_outliers(moneyball2$TEAM_FIELDING_DP)
moneyball2$TEAM_PITCHING_BB_NEW <- treat_outliers(moneyball2$TEAM_PITCHING_BB)
moneyball2$TEAM_PITCHING_H_NEW <- treat_outliers(moneyball2$TEAM_PITCHING_H)
moneyball2$TEAM_PITCHING_HR_NEW <- treat_outliers(moneyball2$TEAM_PITCHING_HR)
moneyball2$TEAM_PITCHING_SO_NEW <- treat_outliers(moneyball2$TEAM_PITCHING_SO)

```


Lets see how the new variables look in boxplots.

```{r}

par(mfrow=c(3,4))

boxplot(moneyball2$TEAM_BATTING_H_NEW,main="TEAM_BATTING_H_NEW")
boxplot(moneyball2$TEAM_BATTING_2B_NEW,main="TEAM_BATTING_2B_NEW")
boxplot(moneyball2$TEAM_BATTING_3B_NEW,main="TEAM_BATTING_3B_NEW")
boxplot(moneyball2$TEAM_BATTING_BB_NEW,main="TEAM_BATTING_BB_NEW")
boxplot(moneyball2$TEAM_BASERUN_SB_NEW,main="TEAM_BASERUN_SB_NEW")
boxplot(moneyball2$TEAM_FIELDING_E_NEW,main="TEAM_FIELDING_E_NEW")
boxplot(moneyball2$TEAM_FIELDING_DP_NEW,main="TEAM_FIELDING_DP_NEW")
boxplot(moneyball2$TEAM_PITCHING_BB_NEW,main="TEAM_PITCHING_BB_NEW")
boxplot(moneyball2$TEAM_PITCHING_H_NEW,main="TEAM_PITCHING_H_NEW")
boxplot(moneyball2$TEAM_PITCHING_HR_NEW,main="TEAM_PITCHING_HR_NEW")
boxplot(moneyball2$TEAM_PITCHING_SO_NEW,main="TEAM_PITCHING_SO_NEW")

```
\
\
\

In the second set, we will use the sin transformation and create the following variables:

TEAM_BATTING_H_SIN
TEAM_BATTING_2B_SIN
TEAM_BATTING_3B_SIN
TEAM_BATTING_BB_SIN
TEAM_BASERUN_SB_SIN
TEAM_FIELDING_E_SIN
TEAM_FIELDING_DP_SIN
TEAM_PITCHING_BB_SIN
TEAM_PITCHING_H_SIN
TEAM_PITCHING_HR_SIN
TEAM_PITCHING_SO_SIN

```{r}

moneyball2$TEAM_BATTING_H_SIN <- sin(moneyball2$TEAM_BATTING_H)
moneyball2$TEAM_BATTING_2B_SIN <- sin(moneyball2$TEAM_BATTING_2B)
moneyball2$TEAM_BATTING_3B_SIN <- sin(moneyball2$TEAM_BATTING_3B)
moneyball2$TEAM_BATTING_BB_SIN <- sin(moneyball2$TEAM_BATTING_BB)
moneyball2$TEAM_BASERUN_SB_SIN <- sin(moneyball2$TEAM_BASERUN_SB)
moneyball2$TEAM_FIELDING_E_SIN <- sin(moneyball2$TEAM_FIELDING_E)
moneyball2$TEAM_FIELDING_DP_SIN <- sin(moneyball2$TEAM_FIELDING_DP)
moneyball2$TEAM_PITCHING_BB_SIN <- sin(moneyball2$TEAM_PITCHING_BB)
moneyball2$TEAM_PITCHING_H_SIN <- sin(moneyball2$TEAM_PITCHING_H)
moneyball2$TEAM_PITCHING_HR_SIN <- sin(moneyball2$TEAM_PITCHING_HR)
moneyball2$TEAM_PITCHING_SO_SIN <- sin(moneyball2$TEAM_PITCHING_SO)

```

\
\

## Missing Values

Next we impute missing values. Since we have handled outliers, we can go ahead and use the mean as impute values. As with outliers, we will go ahead and create new variables for the following:

TEAM_BATTING_SO_NEW

We will re-use the already created new variables for fixing the missing values for the below:

TEAM_PITCHING_SO_NEW 
TEAM_BASERUN_SB_NEW
TEAM_FIELDING_DP_NEW


```{r}

moneyball2$TEAM_BATTING_SO_NEW <- moneyball2$TEAM_BATTING_SO
moneyball2$TEAM_BATTING_SO_NEW[is.na(moneyball2$TEAM_BATTING_SO_NEW)] <- mean(moneyball2$TEAM_BATTING_SO_NEW, na.rm = T) 

moneyball2$TEAM_PITCHING_SO_NEW[is.na(moneyball2$TEAM_PITCHING_SO_NEW)] <- mean(moneyball2$TEAM_PITCHING_SO_NEW, na.rm = T) 
moneyball2$TEAM_BASERUN_SB_NEW[is.na(moneyball2$TEAM_BASERUN_SB_NEW)] <- mean(moneyball2$TEAM_BASERUN_SB_NEW, na.rm = T) 
moneyball2$TEAM_FIELDING_DP_NEW[is.na(moneyball2$TEAM_FIELDING_DP_NEW)] <- mean(moneyball2$TEAM_FIELDING_DP_NEW, na.rm = T) 

```

## Additional Variables 

Lets now create some additional variables that might help us in out analysis. 

### Missing Flags

First we create flag variables to indicate whether TEAM_BATTING_HBP and TEAM_BASERUN_CS and missing. If the value is missing, we code it with 1 and if the value is present we code it with 0.

```{r}

moneyball2$TEAM_BATTING_HBP_Missing <- ifelse(complete.cases(moneyball2$TEAM_BATTING_HBP),1,0)
moneyball2$TEAM_BASERUN_CS_Missing <- ifelse(complete.cases(moneyball2$TEAM_BASERUN_CS),1,0)

```
\
\


###Ratios

\

Next we create some additional variables, that we think may be useful with the prediction. Here we create the following ratios:


```{r}

moneyball2$Hits_R <- moneyball2$TEAM_BATTING_H/moneyball2$TEAM_PITCHING_H
moneyball2$Walks_R <- moneyball2$TEAM_BATTING_BB/moneyball2$TEAM_PITCHING_BB
moneyball2$HomeRuns_R <- moneyball2$TEAM_BATTING_HR/moneyball2$TEAM_PITCHING_HR
moneyball2$Strikeout_R <- moneyball2$TEAM_BATTING_SO/moneyball2$TEAM_PITCHING_SO
```

\
\


###Calculated Variables

\

Finally, we create some calculated variables as below:

1. TEAM_BATTING_EB (Extra Base Hits) = 2B + 3B + HR
2. TEAM_BATTING_1B (Singles by batters) = TEAM_BATTING_H - TEAM_BATTING_EB

```{r}

moneyball2$TEAM_BATTING_EB <- moneyball2$TEAM_BATTING_2B + moneyball2$TEAM_BATTING_3B + moneyball2$TEAM_BATTING_HR

moneyball2$TEAM_BATTING_1B <- moneyball2$TEAM_BATTING_H - moneyball2$TEAM_BATTING_EB
summary(moneyball2)

```

