---
title: "Business Analytics - Final Project"
author:
- Critical Thinking Group 5
- Arindam Barman
- Mohamed Elmoudni
- Shazia Khan
- Kishore Prasad
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

\newpage

# Overview 

The data set contains approximately 41188 obs. of 21 variables. \

This dataset is based on "Bank Marketing" UCI dataset (please check the description at: http://archive.ics.uci.edu/ml/datasets/Bank+Marketing).
   The data is enriched by the addition of five new social and economic features/attributes (national wide indicators from a ~10M population country), published by the Banco de Portugal and publicly available at: https://www.bportugal.pt/estatisticasweb.\
   
The binary classification goal is to predict if the client will subscribe a bank term deposit (variable y).

#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Missing values and Unique Values \


##1.1	Variable identification

First let's display and examine the data dictionary or the data columns as shown in table 1

```{r, echo = FALSE, warning=FALSE, message=FALSE}
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("MASS",character.only = TRUE)) (install.packages("MASS",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("psych",character.only = TRUE)) (install.packages("psych",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
#if (!require("car",character.only = TRUE)) (install.packages("car",dep=TRUE))
if (!require("faraway",character.only = TRUE)) (install.packages("faraway",dep=TRUE))
if (!require("popbio",character.only = TRUE)) (install.packages("popbio",dep=TRUE))
if (!require("gdata",character.only = TRUE)) (install.packages("gdata",dep=TRUE))
if (!require("reshape",character.only = TRUE)) (install.packages("reshape",dep=TRUE))


#install.packages("fancyvrb")

library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
library(dplyr)
library(psych)
library(stringr)
#library(car)
library(faraway)
library(aod)
library(Rcpp)
library(leaps)
library(ISLR)
library(AUC)
library(ROCR)
library(Amelia)
library(popbio)
library(gdata)
library(reshape)
library(gridExtra)

bank_train <- read.table(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional-full.csv",
           sep = ";",
           header = TRUE)

#variables<- read.csv( "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/Variable%20Description.csv")
#kable(variables, caption = "Variable Description") 


# bank_test <-read.table("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional.csv",sep = ";",header = TRUE)
# head(bank_test)


```



##1.2 Missing values and Unique Values



```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

par(mfrow=c(1,1))
#finding missing values
missings<- data.frame(sapply(bank_train,function(x) sum(is.na(x))))
names(missings)[1]<- paste("Missing Values")
kable(missings, caption = "Missing Values")

# missing values graph
#missmap(bank_train, main = "Missing values vs observed")

### finding unique values
uniques<- data.frame(sapply(bank_train, function(x) length(unique(x))))
names(uniques)[1]<- paste("Unique Values")
kable(uniques, caption = "Unique Values")

str(bank_train)

#prop.table(table(bank_train$y))




```


#2	Data Preparation

-Convert Binary to 0 and 1\
-Create dummy variables\
-Data Summary Analysis \
-Correlation of Variables with y\

```{r}
```


##2.1 Convert Binary yes and no to 0 and 1

Now in order to prepare the data for modeling, we need to update Yes = 1 and No = 0. \

```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

bank_train2<-bank_train

#update response variable to binary values of 0 and 1
#levels(bank_train2$y)
levels(bank_train2$y) <- c(0, 1)
bank_train2$y <- as.numeric(levels(bank_train2$y))[bank_train2$y]
str(bank_train2)

```


##2.2 Create dummy variables


```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

#create dummy variables for job values
for(level in unique(bank_train2$job)){
  bank_train2[paste("job", level, sep = "_")] <- ifelse(bank_train2$job == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$job <- NULL
#head(bank_train2)

#create dummy variables for marital values
#levels(bank_train2$marital)
for(level in unique(bank_train2$marital)){
  bank_train2[paste("marital", level, sep = "_")] <- ifelse(bank_train2$marital == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$marital <- NULL
#head(bank_train2)


#education dummy variables primary, secondary tertiary, unknown and illiterate

#education_None
bank_train2$education_illiterate <- as.numeric(ifelse(bank_train2$education == "illiterate", 1, 0))
#table(bank_train2$education_Illiterate)

#education_Unknown
bank_train2$education_unknown <-as.numeric(ifelse(bank_train2$education == "unknown", 1, 0))
#table(bank_train2$education_Unknown)

#education_Primary
bank_train2$education_primary <- as.numeric(ifelse(bank_train2$education == "basic.4y" 
                                        | bank_train2$education == "basic.6y", 1, 0))
#table(bank_train2$education_Primary)

#education_Secondary
bank_train2$education_secondary <- as.numeric(ifelse(bank_train2$education == "basic.9y" 
                                        | bank_train2$education == "high.school", 1, 0))
#table(bank_train2$education_Secondary)

#education_Tertiary
bank_train2$education_tertiary <- as.numeric(ifelse(bank_train2$education == "professional.course" 
                                        | bank_train2$education == "university.degree", 1, 0))
#table(bank_train2$education_Tertiary)

#Delete original catagorical variable
bank_train2$education <- NULL


# contact has 2 levels - 1 variable is required
#levels(bank_train2$contact)
for(level in unique(bank_train2$contact)){
  bank_train2[paste("contact", level, sep = "_")] <- ifelse(bank_train2$contact == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$contact <- NULL

# contact has 2 levels - 1 variable is required
#levels(bank_train2$default)
for(level in unique(bank_train2$default)){
  bank_train2[paste("default", level, sep = "_")] <- ifelse(bank_train2$default == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$default <- NULL


#levels(bank_train2$housing)
for(level in unique(bank_train2$housing)){
  bank_train2[paste("housing", level, sep = "_")] <- ifelse(bank_train2$housing == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$housing <- NULL

#levels(bank_train2$loan)
for(level in unique(bank_train2$loan)){
  bank_train2[paste("loan", level, sep = "_")] <- ifelse(bank_train2$loan == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$loan <- NULL

#levels(bank_train2$poutcome)
for(level in unique(bank_train2$poutcome)){
  bank_train2[paste("poutcome", level, sep = "_")] <- ifelse(bank_train2$poutcome == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$poutcome <- NULL

#levels(bank_train2$month)
for(level in unique(bank_train2$month)){
  bank_train2[paste("month", level, sep = "_")] <- ifelse(bank_train2$month == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$month <- NULL

#levels(bank_train2$day_of_week)
for(level in unique(bank_train2$day_of_week)){
  bank_train2[paste("day_of_week", level, sep = "_")] <- ifelse(bank_train2$day_of_week == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$day_of_week <- NULL

#dummy variable for previous contact yes or no ; 1 or 0 
bank_train2$previous_contact <- as.numeric(ifelse(bank_train2$pdays == 999, 0, 1))
#Delete original numeric variable
bank_train2$pdays <- NULL


#Dummy variables that are k and not required based on correlations
bank_train2$contact_telephone <- NULL
bank_train2$default_unknown <- NULL
bank_train2$marital_married <- NULL
bank_train2$education_secondary <- NULL
bank_train2$poutcome_nonexistent <- NULL
bank_train2$day_of_week_mon <- NULL
bank_train2$month_may <- NULL
bank_train2$`job_blue-collar`<- NULL
bank_train2$housing_no <- NULL




```

##2.3 Data Summary Analysis 

```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

#str(bank_train2)

ds_stats <- psych::describe(bank_train2, skew = TRUE, na.rm = TRUE)
#ds_stats
kable(ds_stats[1:7], caption= "Data Summary")
kable(ds_stats[8:13], caption= "Data Summary (Cont)")

#head(bank_train2)

fun1 <- function(a, y) cor(y, a)
x<-bank_train2[,]
Correlation <- sapply(bank_train2[,], FUN = fun1, bank_train2$y) 



```

##2.4 Correlation of Variables with y 
Now we will produce the correlation table between the independent variables and the dependent variable

```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

Correlation <- sort(Correlation, decreasing = TRUE)
#head(Correlation)
kable(Correlation, caption = "Variable Correlation")

#str(bank_train2)
#str(bank_train)
#summary(bank_train2)

```

##2.5 # Outliers

```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

mdata<- select(bank_train2,  age, previous, duration, campaign, emp.var.rate, cons.price.idx, euribor3m,nr.employed)
mdata2 <- melt(mdata)

# Output the boxplot
p <- ggplot(data = mdata2, aes(x=variable, y=value)) + 
  geom_boxplot() + ggtitle("Outliers Identification")
p + facet_wrap( ~ variable, scales="free", ncol=4)



```

##2.6 Analysis the link function \




```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}

#move y to the last column
bank_train2 <- bank_train2 %>%
  select(-y, everything())

#vars <- names(Correlation)
vars<-names(bank_train2)

par(mfrow=c(2,4))
#Show in the order of Correlation
p = list()
#for (i in 2:ncol(bank_train2)) p[[i]] <- qplot(bank_train2[,i], xlab=names(bank_train2)[[i]])
for (i in 2:ncol(bank_train2)) {
  p[[i]] <- logi.hist.plot(bank_train2[,vars[i]],bank_train2$y,logi.mod = 1, type='hist', boxp=FALSE,col='gray', mainlabel = vars[i])
  
}
#do.call(grid.arrange, p)
#plot(p)
#plot(p[[1]], p[[2]])
#plot (p$your.x.coordinate, p$your.y.coordinate)
#head(bank_train2)


p[[8]] <- logi.hist.plot(bank_train2[,vars[8]],bank_train2$y,logi.mod = 1, type='hist', boxp=FALSE,col='gray', mainlabel = vars[8])

```


##2.7 Interpretation 


### Analysis of the predictor variables:

```{r Analysis on Age variable, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,1))

plot(y~age,bank_train,main="Age vs Response")

```
\
Analysis: from the chart above, it can be seen how the campaign response varies with different age group. It appears that better response was received from young age group between 15-30 and then age group above 60. \
\
\
```{r Analysis on job variable, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,1))

plot(y~job,bank_train,main="Job vs Response")

```
\
Analysis: from the chart above, it can be seen among student and retired personel proportion of positive response is higher though their propostion in overall population is low. Among other profession Admin looks ike having the higher total number of responses\
\
\
```{r Analysis on marital status variable, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,1))

plot(y~marital,bank_train,main="Marital vs Response")

```

\

Responses from single status folks were relatively higher.\
\
\
```{r Analysis on education variable, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,1))

plot(y~education,bank_train,main="Education vs Response")

```
\
Responses from university educated person is relatively higher.\

```{r Analysis on education + Job variable, echo = FALSE, warning=FALSE, message=FALSE}

ggplot(data=bank_train,  aes(x=job, y=education, group=y,
           shape=y, color=y) ) + geom_point() 
  
 
        
```
\
Above diagram shows interaction between variable Job and education with response variable

```{r Analysis on default variable, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,3))

plot(y~default,bank_train,main="default vs Response")
plot(y~loan,bank_train,main="loan vs Response")
plot(y~housing,bank_train,main="housing vs Response")

```
\
With loan and housing variable proportion of responses do not change as shwon from chart


```{r Analysis on loan  housing variable, echo = FALSE, warning=FALSE, message=FALSE}

ggplot(data=bank_train,  aes(x=loan, y=housing, group=y,
           shape=y, color=y) ) + geom_point() 

        
```

\


```{r Analysis on contact type variable, echo = FALSE, warning=FALSE, message=FALSE}

plot(y~contact,bank_train)
        
```
\
Responses from contact of type cellular is higher compare to other types
\

```{r Analysis on number of Contacts  Campaign variable, echo = FALSE, warning=FALSE, message=FALSE}

plot(y~campaign,bank_train)

ggplot(data=bank_train,  aes(x=contact, y=campaign, group=y,
           shape=y, color=y) ) + geom_point() 
  
        
```
\
It can be seen higher number of interactions has reduced the number of positive responses.



```{r Analysis on call variable, echo = FALSE, warning=FALSE, message=FALSE}
plot(table(bank_train$previous,bank_train$y))
      
```
\
Responses from contact of type cellular is higher compare to other types
\


```{r Analysis on call campaign variable, echo = FALSE, warning=FALSE, message=FALSE}
plot(table(bank_train$previous,bank_train$y))
plot(table(bank_train$y,bank_train$poutcome)) 
     
```


```{r Analysis on outcomee  variable, echo = FALSE, warning=FALSE, message=FALSE}

ggplot(data=bank_train,  aes(x=previous, y=poutcome, group=y,
           shape=y, color=y) ) + geom_point() 
  
```




```{r Analysis on number of Consumer index  variable, echo = FALSE, warning=FALSE, message=FALSE}

ggplot(data=bank_train,  aes(x=cons.price.idx, y=cons.conf.idx, group=y,
           shape=y, color=y) ) + geom_point() 
  
```




##2.8 Interaction analysis among variables




```{r Analysis of given variable age and marital status,echo = FALSE, warning=FALSE, message=FALSE}

# Analysis of variables to identify interactions among the variables

coplot(y~age|marital,panel=panel.smooth,bank_train,rows=1)

coplot(y~campaign|job,panel=panel.smooth,bank_train,rows=1)

coplot(y~campaign|education,panel=panel.smooth,bank_train,rows=1)

coplot(y~campaign|housing,panel=panel.smooth,bank_train,rows=1)

coplot(y~campaign|loan,panel=panel.smooth,bank_train,rows=1)

coplot(y~previous|poutcome,panel=panel.smooth,bank_train,rows=1)

cor(bank_train$cons.price.idx,bank_train$cons.conf.idx)

coplot(y~cons.price.idx|cons.conf.idx,panel=panel.smooth,bank_train,rows=1)







```


