---
title: "Business Analytics - Final Project"
author:
- Critical Thinking Group 5
- Arindam Barman
- Mohamed Elmoudni
- Shazia Khan
- Kishore Prasad
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

# Overview 

The data set contains approximately 41188 obs. of 21 variables. \

This dataset is based on "Bank Marketing" UCI dataset (please check the description at: http://archive.ics.uci.edu/ml/datasets/Bank+Marketing).
   The data is enriched by the addition of five new social and economic features/attributes (national wide indicators from a ~10M population country), published by the Banco de Portugal and publicly available at: https://www.bportugal.pt/estatisticasweb.\
   
The binary classification goal is to predict if the client will subscribe a bank term deposit (variable y).

This dependent variable tells whether the client will subscribe a bank term deposit or not. This is a binary variable and as such we will be using a Logistic Regression Model.

#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Understanding predictor variables relationship with response variable
-Missing values and Unique Values \


##1.1	Variable identification 

First let's display and examine the data dictionary or the data columns as shown in table 1

```{r, echo = FALSE, warning=FALSE, message=FALSE}
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("MASS",character.only = TRUE)) (install.packages("MASS",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("psych",character.only = TRUE)) (install.packages("psych",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
#if (!require("car",character.only = TRUE)) (install.packages("car",dep=TRUE))
if (!require("faraway",character.only = TRUE)) (install.packages("faraway",dep=TRUE))
if (!require("popbio",character.only = TRUE)) (install.packages("popbio",dep=TRUE))
if (!require("gdata",character.only = TRUE)) (install.packages("gdata",dep=TRUE))
if (!require("reshape",character.only = TRUE)) (install.packages("reshape",dep=TRUE))


#install.packages("fancyvrb")

library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
library(dplyr)
library(psych)
library(stringr)
#library(car)
library(faraway)
library(aod)
library(Rcpp)
library(leaps)
library(ISLR)
library(AUC)
library(ROCR)
library(Amelia)
library(popbio)
library(gdata)
library(reshape)
library(gridExtra)

bank_train <- read.table(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional-full.csv",
           sep = ";",
           header = TRUE)

bank_test <-read.table(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional.csv",
           sep = ";",
           header = TRUE)

```

We notice that the variables are numerical, categorical and binary. The responce variable y is binary.

Based on the original dataset, our predictor input has 21 variables. And our response variable is 1 variable called y. 

Binomial Logistic regression is the appropriate regression analysis to conduct when the dependent variable is dichotomous (binary).  Like all regression analyses, the logistic regression is a predictive analysis.  Logistic regression is used to describe data and to explain the relationship between one dependent binary variable and one or more metric (interval or ratio scale) independent variables.


##1.2 Preliminary Data Analysis

##1.2.1 Analysis of Predictor variable 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

variable_analysis<- read.csv(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/Variable%20Analysis.csv")
kable(variable_analysis, caption = "Variable Analysis") 

```


##1.2.2 Missing values 

We see that there are no missing values in our dataset as shown in table 2 and graph format.
The unique values are given in the table 

```{r, echo = FALSE, warning=FALSE, message=FALSE}



```

##1.2.3 Proportion of Response Variables 

```{r, echo = FALSE, warning=FALSE, message=FALSE}



```



#2	Data Preparation

-Convert Binary to 0 and 1\
-Create dummy variables\
-Data Summary Analysis \
-Correlation of Variables with y\


##2.1 Convert Binary yes and no to 0 and 1

Now in order to prepare the data for modeling, we need to update Yes = 1 and No = 0. \

```{r, echo = FALSE, warning=FALSE, message=FALSE}

bank_train2<-bank_train

#update response variable to binary values of 0 and 1
#levels(bank_train2$y)
levels(bank_train2$y) <- c(0, 1)
bank_train2$y <- as.numeric(levels(bank_train2$y))[bank_train2$y]
#str(bank_train2)

```

##2.2 Create dummy variables

Now we need to create dummy variables to find out the relationship between y variables and dependent variables, for all categorical variables.\
```{r, echo = FALSE, warning=FALSE, message=FALSE}

# age is numeric 

#create dummy variables for job values
for(level in unique(bank_train2$job)){
  bank_train2[paste("job", level, sep = "_")] <- ifelse(bank_train2$job == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$job <- NULL
#head(bank_train2)

#create dummy variables for marital values
#levels(bank_train2$marital)
for(level in unique(bank_train2$marital)){
  bank_train2[paste("marital", level, sep = "_")] <- ifelse(bank_train2$marital == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$marital <- NULL
#head(bank_train2)

#--------------------------------------------------------------------------
#education dummy variables primary, secondary tertiary, unknown and illiterate

#education_None
bank_train2$education_illiterate <- as.numeric(ifelse(bank_train2$education == "illiterate", 1, 0))
#table(bank_train2$education_Illiterate)

#education_Unknown
bank_train2$education_unknown <-as.numeric(ifelse(bank_train2$education == "unknown", 1, 0))
#table(bank_train2$education_Unknown)

#education_Primary
bank_train2$education_primary <- as.numeric(ifelse(bank_train2$education == "basic.4y" 
                                        | bank_train2$education == "basic.6y", 1, 0))
#table(bank_train2$education_Primary)

#education_Secondary
bank_train2$education_secondary <- as.numeric(ifelse(bank_train2$education == "basic.9y" 
                                        | bank_train2$education == "high.school", 1, 0))
#table(bank_train2$education_Secondary)

#education_Tertiary
bank_train2$education_tertiary <- as.numeric(ifelse(bank_train2$education == "professional.course" 
                                        | bank_train2$education == "university.degree", 1, 0))
#table(bank_train2$education_Tertiary)

#Delete original catagorical variable
bank_train2$education <- NULL


# contact has 2 levels - 1 variable is required
#levels(bank_train2$default)
for(level in unique(bank_train2$default)){
  bank_train2[paste("default", level, sep = "_")] <- ifelse(bank_train2$default == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$default <- NULL


#levels(bank_train2$housing)
for(level in unique(bank_train2$housing)){
  bank_train2[paste("housing", level, sep = "_")] <- ifelse(bank_train2$housing == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$housing <- NULL

#levels(bank_train2$loan)
for(level in unique(bank_train2$loan)){
  bank_train2[paste("loan", level, sep = "_")] <- ifelse(bank_train2$loan == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$loan <- NULL

# contact has 2 levels - 1 variable is required
#levels(bank_train2$contact)
for(level in unique(bank_train2$contact)){
  bank_train2[paste("contact", level, sep = "_")] <- ifelse(bank_train2$contact == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$contact <- NULL

#levels(bank_train2$month)
for(level in unique(bank_train2$month)){
  bank_train2[paste("month", level, sep = "_")] <- ifelse(bank_train2$month == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$month <- NULL

#levels(bank_train2$day_of_week)
for(level in unique(bank_train2$day_of_week)){
  bank_train2[paste("day_of_week", level, sep = "_")] <- ifelse(bank_train2$day_of_week == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$day_of_week <- NULL

#duration is numeric
#campaign is numeric

#pdays is numeric
#dummy variable for pdays -previous contact yes or no ; 1 or 0  when 999
bank_train2$previous_contact <- as.numeric(ifelse(bank_train2$pdays == 999, 0, 1))

#previous is numeric

#levels(bank_train2$poutcome)
for(level in unique(bank_train2$poutcome)){
  bank_train2[paste("poutcome", level, sep = "_")] <- ifelse(bank_train2$poutcome == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$poutcome <- NULL

# emp.var.rate is numeric
# cons.price.idx is numeric
# cons.conf.idx  is numeric
# euribor3m is numeric
# nr.employed is numeric


```

##2.3 Data Summary Analysis with Dummy variables


##2.4 Correlation between Response and Predictor of Variables
Now we will produce the correlation table between the independent variables and the dependent variable



##2.5 # Outliers

```{r, echo = FALSE, warning=FALSE, message=FALSE}

mdata<- select(bank_train2,  age, previous, duration, campaign, emp.var.rate, cons.price.idx, euribor3m,nr.employed)
mdata2 <- melt(mdata)

# Output the boxplot
p <- ggplot(data = mdata2, aes(x=variable, y=value)) + 
  geom_boxplot() + ggtitle("Outliers Identification")
p + facet_wrap( ~ variable, scales="free", ncol=4)
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}

bank_train2 <- bank_train2 %>%
  select(-y, everything())

fun1 <- function(a, y) cor(y, a)
x<-bank_train2[,]
Correlation <- sapply(x, FUN = fun1, y=bank_train2$y) 


Correlation <- sort(Correlation, decreasing = TRUE)

vars <- names(Correlation)


par(mfrow=c(2,4))
for (i in 2:ncol(bank_train2)) {
  hist(bank_train2[,vars[i]], main = vars[i], xlab = "")
}

```

##2.6 Analysis the link function \

In this section, we will investigate how our initial data aligns with a typical logistic model plot.

Recall the Logistic Regression is part of a larger class of algorithms known as Generalized Linear Model (glm). The fundamental equation of generalized linear model is:

$g(E(y)) = a+ Bx_1+B_2x_2+ B_3x_3+...$

where, g() is the link function, E(y) is the expectation of target variable and $B_0 + B_1x_1 + B_2x_2+B_3x_3$ is the linear predictor ( $B_0,B_1,B_2, B_3$ to be predicted). The role of link function is to 'link' the expectation of y to linear predictor.

In logistic regression, we are only concerned about the probability of outcome dependent variable ( success or failure). As described above, g() is the link function. This function is established using two things: Probability of Success (p) and Probability of Failure (1-p). p should meet following criteria: It must always be positive (since p >= 0) It must always be less than equals to 1 (since p <= 1).

Now let's investigate how our initial data model aligns with the above criteria. In other words, we will plot regression model plots for each variable and compare it to a typical logistic model plot:


The main objective in the transformations is to achieve linear relationships with the dependent variable (or, really, with its logit).

```{r, echo = FALSE, warning=FALSE, message=FALSE}

#move y to the last column


par(mfrow=c(2,4))
#Show in the order of Correlation
p = list()
#for (i in 2:ncol(bank_train2)) p[[i]] <- qplot(bank_train2[,i], xlab=names(bank_train2)[[i]])
for (i in 2:ncol(bank_train2)) {
  p[[i]] <- logi.hist.plot(bank_train2[,vars[i]],bank_train2$y,logi.mod = 1, type='hist', boxp=FALSE,col='gray', 
                           mainlabel = vars[i])
}
#do.call(grid.arrange, p)
#plot(p)
#plot(p[[1]], p[[2]])
#plot (p$your.x.coordinate, p$your.y.coordinate)
#head(bank_train2)

 
```

#3 Build Models

In this section, we will create 3 models. Aside from using original and transformed data, we will also be using different methods and functions such as Linear Discriminant Analysis, step function, and logit function to enhance our models. 

Below is our model definition: 

-Model 1- This model will be created using all the variables in train data set with logit function GLM. 

-Model 2: This model step function will be used to enhance the model 1. 

-Model 3- This model will be created using calssification and regression tree. 

## 3.1 Model 1

Taking the treated data and splitting into 80/20 to train model and validate the data.

```{r, echo = FALSE, warning=FALSE, message=FALSE}

smp_size <- floor(0.80 * nrow(bank_train2))

## set the seed to make your partition reproductible
set.seed(123)

train_index <- sample(seq_len(nrow(bank_train2)), size = smp_size)

DS_TARGET_FLAG_TRAIN<- bank_train2[train_index, ]
DS_TARGET_FLAG_VALID <- bank_train2[-train_index, ]

```


```{r, echo = FALSE, warning=FALSE, message=FALSE,eval=FALSE}


model1 <- glm(y ~.,family=binomial(link='logit'),data=DS_TARGET_FLAG_TRAIN)
summary(model1)
anova(model1, test="Chisq")

library(pscl)
#the McFadden R^2 index can be used to assess the model fit.
pR2(model1)

fitted.results <- predict(model1,newdata=DS_TARGET_FLAG_VALID,type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)

misClasificError <- mean(fitted.results != DS_TARGET_FLAG_VALID$y)
print(paste('Accuracy',1-misClasificError))


```

## 3.2 Model 2

## 3.3 Model 3

#4 Evaluate Models

#5 Select Models

#6 Appendix



#6.1  Data Analysis details

#6.1.1  Variable Description

```{r, echo = FALSE, warning=FALSE, message=FALSE}

variables<- read.csv(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/Variable%20Description.csv")
kable(variables, caption = "Variable Description") 

```
\

#6.1.2  Predictor and Response variable Association
```{r, echo = FALSE, warning=FALSE, message=FALSE}


#round(prop.table(table(bank_train$y, bank_train$age),2)*100,2)  #student, retired
ggplot(bank_train, aes(age)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$job),2)*100,2)  #student, retired
ggplot(bank_train, aes(job)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$marital),2)*100,2) #unknown
ggplot(bank_train, aes(marital)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$education),2)*100,2) #illeterate
ggplot(bank_train, aes(education)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$default),2)*100,2) #no
ggplot(bank_train, aes(default)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$housing),2)*100,2) #no
ggplot(bank_train, aes(housing)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$loan),2)*100,2) #no
ggplot(bank_train, aes(loan)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$contact),2)*100,2) #cellular
ggplot(bank_train, aes(contact)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$month),2)*100,2) #march, dec, sep, oct
ggplot(bank_train, aes(month)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$day_of_week),2)*100,2) #Thursday
ggplot(bank_train, aes(day_of_week)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$duration),2)*100,2) #increases with every contact upto 5
ggplot(bank_train, aes(duration)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+ xlim(0, 2200)

#round(prop.table(table(bank_train$y, bank_train$campaign),2)*100,2) #increases with every contact upto 5
ggplot(bank_train, aes(campaign)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$pdays),2)*100,2) #increases with every contact upto 5
ggplot(bank_train, aes(pdays)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

#round(prop.table(table(bank_train$y, bank_train$previous),2)*100,2) #increases with every contact upto 5
ggplot(bank_train, aes(previous)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

  #round(prop.table(table(bank_train$y, bank_train$poutcome),2)*100,2) #success
ggplot(bank_train, aes(poutcome)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

 #round(prop.table(table(bank_train$y, bank_train$emp.var.rate),2)*100,2) #success
ggplot(bank_train, aes(emp.var.rate)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

 #round(prop.table(table(bank_train$y, bank_train$cons.price.idx),2)*100,2) #success
ggplot(bank_train, aes(cons.price.idx)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

 #round(prop.table(table(bank_train$y, bank_train$cons.conf.idx),2)*100,2) #success
ggplot(bank_train, aes(cons.conf.idx)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

 #round(prop.table(table(bank_train$y, bank_train$euribor3m),2)*100,2) #success
ggplot(bank_train, aes(euribor3m)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

 #round(prop.table(table(bank_train$y, bank_train$nr.employed),2)*100,2) #success
ggplot(bank_train, aes(nr.employed)) + geom_bar(aes(fill = y), position = "fill")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))



```

#6.1.3  Unique Value & Missing value

We see that there are no missing values in our dataset as shown in table 2 and graph format.
The unique values are given in the table 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(1,1))
#finding missing values
missings<- data.frame(sapply(bank_train,function(x) sum(is.na(x))))
names(missings)[1]<- paste("Missing Values")
kable(missings, caption = "Missing Values")

# missing values graph
#missmap(bank_train, main = "Missing values vs observed")

### finding unique values
uniques<- data.frame(sapply(bank_train, function(x) length(unique(x))))
names(uniques)[1]<- paste("Unique Values")
kable(uniques, caption = "Unique Values")

```


##6.1.3.4 Data Summary post conversion 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

#str(bank_train2)

ds_stats <- psych::describe(bank_train2, skew = TRUE, na.rm = TRUE)
#ds_stats
kable(ds_stats[1:7], caption= "Data Summary")
kable(ds_stats[8:13], caption= "Data Summary (Cont)")

#head(bank_train2)

fun1 <- function(a, y) cor(y, a)
x<-bank_train2[,]
Correlation <- sapply(x, FUN = fun1, y=bank_train2$y) 


Correlation <- sort(Correlation, decreasing = TRUE)
#head(Correlation)
kable(Correlation, caption = "Variable Correlation")

#str(bank_train2)
#str(bank_train)
#summary(bank_train2)

```

