---
title: "Business Analytics - Final Project"
author:
- Critical Thinking Group 5
- Arindam Barman
- Mohamed Elmoudni
- Shazia Khan
- Kishore Prasad
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

\newpage

# Overview 

The data set contains approximately 41188 obs. of 21 variables. \

This dataset is based on "Bank Marketing" UCI dataset (please check the description at: http://archive.ics.uci.edu/ml/datasets/Bank+Marketing).
   The data is enriched by the addition of five new social and economic features/attributes (national wide indicators from a ~10M population country), published by the Banco de Portugal and publicly available at: https://www.bportugal.pt/estatisticasweb.\
   
The binary classification goal is to predict if the client will subscribe a bank term deposit (variable y).

#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Missing values and Unique Values \


##1.1	Variable identification

First let's display and examine the data dictionary or the data columns as shown in table 1

```{r, echo = FALSE, warning=FALSE, message=FALSE}
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("MASS",character.only = TRUE)) (install.packages("MASS",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("psych",character.only = TRUE)) (install.packages("psych",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
#if (!require("car",character.only = TRUE)) (install.packages("car",dep=TRUE))
if (!require("faraway",character.only = TRUE)) (install.packages("faraway",dep=TRUE))
if (!require("popbio",character.only = TRUE)) (install.packages("popbio",dep=TRUE))
if (!require("gdata",character.only = TRUE)) (install.packages("gdata",dep=TRUE))
if (!require("reshape",character.only = TRUE)) (install.packages("reshape",dep=TRUE))


#install.packages("fancyvrb")

library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
library(dplyr)
library(psych)
library(stringr)
#library(car)
library(faraway)
library(aod)
library(Rcpp)
library(leaps)
library(ISLR)
library(AUC)
library(ROCR)
library(Amelia)
library(popbio)
library(gdata)
library(reshape)
library(gridExtra)

bank_train <- read.table(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional-full.csv",
           sep = ";",
           header = TRUE)

variables<- read.csv(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/Variable%20Description.csv")
kable(variables, caption = "Variable Description") 


bank_test <-read.table(
  "https://raw.githubusercontent.com/kishkp/data621-ctg5/master/Final%20Project/bank-additional.csv",
           sep = ";",
           header = TRUE)
# head(bank_test)


```

We notice that the variables are numerical, categorical and binary. The responce variable y is binary.

Based on the original dataset, our predictor input has 21 variables. And our response variable is 1 variable called y. 

Binomial Logistic regression is the appropriate regression analysis to conduct when the dependent variable is dichotomous (binary).  Like all regression analyses, the logistic regression is a predictive analysis.  Logistic regression is used to describe data and to explain the relationship between one dependent binary variable and one or more metric (interval or ratio scale) independent variables.


##1.2 Missing values and Unique Values

We see that there are no missing values in our dataset as shown in table 2 and graph format.
The unique values are given in the table 3.

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(1,1))
#finding missing values
missings<- data.frame(sapply(bank_train,function(x) sum(is.na(x))))
names(missings)[1]<- paste("Missing Values")
kable(missings, caption = "Missing Values")

# missing values graph
#missmap(bank_train, main = "Missing values vs observed")

### finding unique values
uniques<- data.frame(sapply(bank_train, function(x) length(unique(x))))
names(uniques)[1]<- paste("Unique Values")
kable(uniques, caption = "Unique Values")

str(bank_train)

#prop.table(table(bank_train$y))




```



#2	Data Preparation

-Convert Binary to 0 and 1\
-Create dummy variables\
-Data Summary Analysis \
-Correlation of Variables with y\

```{r}
```


##2.1 Convert Binary yes and no to 0 and 1

Now in order to prepare the data for modeling, we need to update Yes = 1 and No = 0. \

```{r, echo = FALSE, warning=FALSE, message=FALSE}

bank_train2<-bank_train

#update response variable to binary values of 0 and 1
#levels(bank_train2$y)
levels(bank_train2$y) <- c(0, 1)
bank_train2$y <- as.numeric(levels(bank_train2$y))[bank_train2$y]
str(bank_train2)

```


##2.2 Create dummy variables

Now we need to create dummy variables to find out the relationship between y variables and dependent variables, for all categorical variables.\
```{r, echo = FALSE, warning=FALSE, message=FALSE}

#create dummy variables for job values
for(level in unique(bank_train2$job)){
  bank_train2[paste("job", level, sep = "_")] <- ifelse(bank_train2$job == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$job <- NULL
#head(bank_train2)

#create dummy variables for marital values
#levels(bank_train2$marital)
for(level in unique(bank_train2$marital)){
  bank_train2[paste("marital", level, sep = "_")] <- ifelse(bank_train2$marital == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$marital <- NULL
#head(bank_train2)


#education dummy variables primary, secondary tertiary, unknown and illiterate

#education_None
bank_train2$education_illiterate <- as.numeric(ifelse(bank_train2$education == "illiterate", 1, 0))
#table(bank_train2$education_Illiterate)

#education_Unknown
bank_train2$education_unknown <-as.numeric(ifelse(bank_train2$education == "unknown", 1, 0))
#table(bank_train2$education_Unknown)

#education_Primary
bank_train2$education_primary <- as.numeric(ifelse(bank_train2$education == "basic.4y" 
                                        | bank_train2$education == "basic.6y", 1, 0))
#table(bank_train2$education_Primary)

#education_Secondary
bank_train2$education_secondary <- as.numeric(ifelse(bank_train2$education == "basic.9y" 
                                        | bank_train2$education == "high.school", 1, 0))
#table(bank_train2$education_Secondary)

#education_Tertiary
bank_train2$education_tertiary <- as.numeric(ifelse(bank_train2$education == "professional.course" 
                                        | bank_train2$education == "university.degree", 1, 0))
#table(bank_train2$education_Tertiary)

#Delete original catagorical variable
bank_train2$education <- NULL


# contact has 2 levels - 1 variable is required
#levels(bank_train2$contact)
for(level in unique(bank_train2$contact)){
  bank_train2[paste("contact", level, sep = "_")] <- ifelse(bank_train2$contact == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$contact <- NULL

# contact has 2 levels - 1 variable is required
#levels(bank_train2$default)
for(level in unique(bank_train2$default)){
  bank_train2[paste("default", level, sep = "_")] <- ifelse(bank_train2$default == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$default <- NULL


#levels(bank_train2$housing)
for(level in unique(bank_train2$housing)){
  bank_train2[paste("housing", level, sep = "_")] <- ifelse(bank_train2$housing == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$housing <- NULL

#levels(bank_train2$loan)
for(level in unique(bank_train2$loan)){
  bank_train2[paste("loan", level, sep = "_")] <- ifelse(bank_train2$loan == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$loan <- NULL

#levels(bank_train2$poutcome)
for(level in unique(bank_train2$poutcome)){
  bank_train2[paste("poutcome", level, sep = "_")] <- ifelse(bank_train2$poutcome == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$poutcome <- NULL

#levels(bank_train2$month)
for(level in unique(bank_train2$month)){
  bank_train2[paste("month", level, sep = "_")] <- ifelse(bank_train2$month == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$month <- NULL

#levels(bank_train2$day_of_week)
for(level in unique(bank_train2$day_of_week)){
  bank_train2[paste("day_of_week", level, sep = "_")] <- ifelse(bank_train2$day_of_week == level, 1, 0)
}
#Delete original catagorical variable
bank_train2$day_of_week <- NULL

#dummy variable for previous contact yes or no ; 1 or 0 
bank_train2$previous_contact <- as.numeric(ifelse(bank_train2$pdays == 999, 0, 1))
#Delete original numeric variable
bank_train2$pdays <- NULL


#Dummy variables that are k and not required based on correlations
bank_train2$contact_telephone <- NULL
bank_train2$default_unknown <- NULL
bank_train2$marital_married <- NULL
bank_train2$education_secondary <- NULL
bank_train2$poutcome_nonexistent <- NULL
bank_train2$day_of_week_mon <- NULL
bank_train2$month_may <- NULL
bank_train2$`job_blue-collar`<- NULL
bank_train2$housing_no <- NULL




```

##2.3 Data Summary Analysis 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

#str(bank_train2)

ds_stats <- psych::describe(bank_train2, skew = TRUE, na.rm = TRUE)
#ds_stats
kable(ds_stats[1:7], caption= "Data Summary")
kable(ds_stats[8:13], caption= "Data Summary (Cont)")

#head(bank_train2)

fun1 <- function(a, y) cor(y, a)
x<-bank_train2[,]
Correlation <- sapply(x, FUN = fun1, y=bank_train2$y) 

```

##2.4 Correlation of Variables with y 
Now we will produce the correlation table between the independent variables and the dependent variable

```{r, echo = FALSE, warning=FALSE, message=FALSE}

Correlation <- sort(Correlation, decreasing = TRUE)
#head(Correlation)
kable(Correlation, caption = "Variable Correlation")

#str(bank_train2)
#str(bank_train)
#summary(bank_train2)

```

##2.5 # Outliers

```{r, echo = FALSE, warning=FALSE, message=FALSE}

mdata<- select(bank_train2,  age, previous, duration, campaign, emp.var.rate, cons.price.idx, euribor3m,nr.employed)
mdata2 <- melt(mdata)

# Output the boxplot
p <- ggplot(data = mdata2, aes(x=variable, y=value)) + 
  geom_boxplot() + ggtitle("Outliers Identification")
p + facet_wrap( ~ variable, scales="free", ncol=4)



```

##2.6 Analysis the link function \

In this section, we will investigate how our initial data aligns with a typical logistic model plot.

Recall the Logistic Regression is part of a larger class of algorithms known as Generalized Linear Model (glm). The fundamental equation of generalized linear model is:

$g(E(y)) = a+ Bx_1+B_2x_2+ B_3x_3+...$

where, g() is the link function, E(y) is the expectation of target variable and $B_0 + B_1x_1 + B_2x_2+B_3x_3$ is the linear predictor ( $B_0,B_1,B_2, B_3$ to be predicted). The role of link function is to 'link' the expectation of y to linear predictor.

In logistic regression, we are only concerned about the probability of outcome dependent variable ( success or failure). As described above, g() is the link function. This function is established using two things: Probability of Success (p) and Probability of Failure (1-p). p should meet following criteria: It must always be positive (since p >= 0) It must always be less than equals to 1 (since p <= 1).

Now let's investigate how our initial data model aligns with the above criteria. In other words, we will plot regression model plots for each variable and compare it to a typical logistic model plot:


```{r, echo = FALSE, warning=FALSE, message=FALSE}

#move y to the last column
bank_train2 <- bank_train2 %>%
  select(-y, everything())

vars <- names(Correlation)

par(mfrow=c(2,4))
#Show in the order of Correlation
p = list()
#for (i in 2:ncol(bank_train2)) p[[i]] <- qplot(bank_train2[,i], xlab=names(bank_train2)[[i]])
for (i in 2:ncol(bank_train2)) {
  p[[i]] <- logi.hist.plot(bank_train2[,vars[i]],bank_train2$y,logi.mod = 1, type='hist', boxp=FALSE,col='gray', 
                           mainlabel = vars[i])
  
}
#do.call(grid.arrange, p)
#plot(p)
#plot(p[[1]], p[[2]])
#plot (p$your.x.coordinate, p$your.y.coordinate)
#head(bank_train2)

```


##2.7 Interpretation 

You can see that the probability of y increases as age increases, duration of the call increses


You can see that the probability of crashing decreases as we get closer to the "1" classification for the URBANICITY_Highly.Rural..Rural, PARENT1_No, REVOKED_No, HOME_VAL, CAR_USE_Private, CAR_TYPE_Minivan, MSTATUS_Yes, JOB_Manager, AGE, CAR_AGE, TIF, EDUCATION_Masters, YOJ, EDUCATION_PhD, JOB_Lawyer, JOB_Doctor, EDUCATION_Bachelors, JOB_Professional, INCOME, SEX_M, RED_CAR_yes, CAR_TYPE_Panel.Truck variables.  \
