---
title: "Home Work Assignment - 03"
author: "Critical Thinking Group 5"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

\newpage

# Overview 

The data set contains approximately 466 records and 14 variables. Each record has information on crime for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime rate is above the median crime rate (1) or not (0).

The objective is to build a binary logistic regression model on the training data set to predict whether the neighborhood will be at risk for high crime levels. In addition, we will provide classifications and probabilities for the evaluation data set using the binary logistic regression model. 

To attain our objective, we will be following the below best practice steps and guidelines:
1 -Data Exploration
2 -Data Preparation
3 -Build Models
4 -Select Models


#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Variable Relationships \
-Data summary analysis \
-Outliers and Missing Values Identification

##1.1	Variable identification

First let's display and examine the data dictionary or the data columns as shown in table 1


```{r, echo = FALSE, warning=FALSE, message=FALSE}

# changed lib path... 

if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("MASS",character.only = TRUE)) (install.packages("MASS",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("psych",character.only = TRUE)) (install.packages("psych",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("car",character.only = TRUE)) (install.packages("car",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("faraway",character.only = TRUE)) (install.packages("faraway",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("aod",character.only = TRUE)) (install.packages("aod",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("ISLR",character.only = TRUE)) (install.packages("ISLR",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("AUC",character.only = TRUE)) (install.packages("AUC",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("ROCR",character.only = TRUE)) (install.packages("ROCR",repos = "http://cran.us.r-project.org", dep=TRUE))
if (!require("leaps",character.only = TRUE)) (install.packages("leaps",repos = "http://cran.us.r-project.org", dep=TRUE))

library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
library(dplyr)
library(psych)
library(stringr)
#library(car)
library(faraway)
library(aod)
library(Rcpp)
library(leaps)
library(ISLR)
library(AUC)
library(ROCR)

city_crime_train_full <- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW3/crime-training-data.csv")

variables<- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW3/variable.csv")
kable(variables, caption = "Variable Description")

#summary(city_crime_train_full)


city_crime_test <- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW3/crime-evaluation-data.csv")

#str(city_crime_test)

```


We notice that all variables are numeric except for two variables: 
the response variable "target" which is binary and the  predictor variable "chas" which 
is a dummy binary variable indicating whether the suburb borders the Charles River (1) or not (0).

Based on the original dataset, our predictor input is made of 13 variables. And our response variable is one variable called target.



```{r}
###finding missing values
missings<- sapply(city_crime_train_full,function(x) sum(is.na(x)))
kable(missings, caption = "Missing Values")
### finding unique values
uniques<- sapply(city_crime_train_full, function(x) length(unique(x)))
kable(uniques, caption = "Unique Values")
```


visually checking missing values 

```{r}
library(Amelia)
missmap(city_crime_train_full, main = "Missing values vs observed")

```


 model 
 
 
```{r}

smp_size <- floor(0.80 * nrow(city_crime_train_full))
## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(city_crime_train_full)), size = smp_size)
city_crime_train<- city_crime_train_full[train_ind, ]
train_test <- city_crime_train_full[-train_ind, ]

model <- glm(target ~.,family=binomial(link='logit'),data=city_crime_train)

summary(model)

```

##### model interpretation

Now we can analyze the fitting and interpret what the model is telling us.
First of all, we can see that indus,chas,rm,age,black, and lstat are not statistically significant. 

As for the statistically significant variables, nox has 
the lowest p-value suggesting a strong association of the nox of 
the person with the probability of being above target.

################# Mychanges


Split the full train data set into train and test to validate the model performance\

1. Split the data 80% train and 20% for model validation

```{r split train data to train and validation, echo=FALSE, eval=TRUE}


smp_size <- floor(0.80 * nrow(city_crime_train_full))

## set the seed to make your partition reproductible
set.seed(123)

train_ind <- sample(seq_len(nrow(city_crime_train_full)), size = smp_size)

city_crime_train<- city_crime_train_full[train_ind, ]
train_test <- city_crime_train_full[-train_ind, ]

```

#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Variable Relationships \
-Data summary analysis \
-Outliers and Missing Values Identification

##1.1	Variable identification

First let's display and examine the data dictionary or the data columns as shown in table 1 and proportion of success and failure cases in target variable. \


```{r, echo = FALSE, warning=FALSE, message=FALSE}

summary(city_crime_train)

par(mfrow=c(1,1))

pairs(city_crime_train,col=city_crime_train$target)

# % of cities marked as crime prone

prop.table(table(city_crime_train$target))

```
\


\newpage

##1.2 Data Summary Analysis


In this section, we will create summary data to better understand the initial relationship variables have with our dependent variable using correlation, central tendency, and dispersion As shown in table 2.  



```{r, echo = FALSE, warning=FALSE, message=FALSE}
ds_stats <- psych::describe(city_crime_train, skew = TRUE, na.rm = TRUE)
ds_stats


x<-colnames(city_crime_train)
fun <- function(x) sum(!complete.cases(x))
Missing <- sapply(city_crime_train[x], FUN = fun) 
Missing


fun1 <- function(x, y) cor(y, x)
Correlation <- sapply(city_crime_train[x], FUN = fun1, y=city_crime_train$target) 

kable(data.frame(Correlation), caption = "Correlation between target and predictor variable")
```

It is clear from the table that most of the variables are having storng correlation with the target variable. \




##1.3	Outliers and Missing Values Identification


In this section we look at boxplots to determine the outliers in variables and decide on whether to act on the outliers. 

Lets do some univariate analysis. We will look at the Histogram and Boxplot for each variable to detect outliers if any and treat it accordingly.\




```{r, echo = FALSE, warning=FALSE, message=FALSE}
show_charts <- function(x, ...) {
    
    par(mfrow=c(2,3))
    
    xlabel <- unlist(str_split(deparse(substitute(x)), pattern = "\\$"))[2]
#    ylabel <- unlist(str_split(deparse(substitute(y)), pattern = "\\$"))[2]
    
    hist(x,main=xlabel)
    boxplot(x,main=xlabel)

    y<-log(x)
    boxplot(y,main='log transform')
    y<-sqrt(x)
    boxplot(y,main='sqrt transform')
    y<-sin(x)
    boxplot(y,main='sin transform')
    y<-(x)^(1/9)
    boxplot(y,main='ninth transform')
   # plot(x, target)
}

attach(city_crime_train)

```

Analysis of variable zn:proportion of residential land zoned for large lots 

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.cap='zn Transformation'}
show_charts(zn)

```
For zn, we can see that there are large number of values with 0. ninth transformation seem better for this variable..(1)

\newpage
***Please note that we have created similar figures to figure 1 above for each remaining variable. 
However, we hid the remaining figures for ease of streamlining the report as they have similar shapes. However, we have drawn the below observations from each remaining figure. \



```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(indus)


```

For indus, we can see that there is a spike toward right side of he distribution. Looking at the sqrt transformation it appears that distribution is close to normal and having two peaks after transformation.  
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(nox)

```
\
For nox, there is a long right tail.

\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(rm)


```

\
For rm, there are some outliers as we can see from box plot. This variable will need some transformation to handle the outliers.
\

```{r, echo = FALSE, warning=FALSE, message=FALSE,fig.show ='hide'}

show_charts(age)
```

age of the building variable is skewed heavily towards right side. We will need some transformation for this variable and looks sin transformation is best option for this case  
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(dis)

```

For this variable dis, there are some outliers which needs transformation to handle those outliers. log transformation looks best suited for this scenario.
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(rad)


```

For rad variable distribution is not uniform as seen from the chart and will need transformation.

\
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(tax)


```

For tax variable is not uniformly distributed but there is no outlier for this variable.
\
\


\
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(ptratio)

```

For pratio has right aligned peak but no outliers are there in data set.
\
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(black)

```


\
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(lstat)

```
The variable lstat has long right tail and lef skewed

\
\

```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.show ='hide'}

show_charts(medv)

```





#2. Data Preparation 

Now that we have completed the preliminary analysis, we will be cleaning and consolidating data into one dataset for use in analysis and modeling. We will be puring the belwo steps as guidlines: \
- Outliers treatment \
- Missing values treatment \
- Data transformation \



##2.1 Outliers treatment

For outliers, we will create 2 sets of variables. 

The first set uses the capping method. In this method, we will replace all outliers that lie outside the 1.5 times of IQR limits. We will cap it by replacing those observations less than the lower limit with the value of 5th %ile and those that lie above the upper limit with the value of 95th %ile. 

Accordingly we create the following new variables while retaining the original variables. 

city_crime_train$tax\
city_crime_train$medv\
city_crime_train$lstat\


```{r, echo = FALSE, warning=FALSE, message=FALSE}
# function for removing outliers - http://r-statistics.co/Outlier-Treatment-With-R.html

treat_outliers <- function(x) {
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
 
return(x)
}

city_crime_train_mod<-city_crime_train
train_test_mod<-train_test

city_crime_train_mod$tax_new <- treat_outliers(city_crime_train$tax)
city_crime_train_mod$medv_new <- treat_outliers(city_crime_train$medv)
city_crime_train_mod$lstat_new <- treat_outliers(city_crime_train$lstat)


train_test_mod$tax_new <- treat_outliers(train_test$tax)
train_test_mod$medv_new <- treat_outliers(train_test$medv)
train_test_mod$lstat_new <- treat_outliers(train_test$lstat)

```


Lets see how the new variables look in boxplots.

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(1,3))

boxplot(city_crime_train_mod$tax_new,main="tax_new")
boxplot(city_crime_train_mod$medv_new,main="medv_new")
boxplot(city_crime_train_mod$lstat_new,main="lstat_new")

```


In the second set, we will use the sin transformation and create the following variables:

city_crime_train_mod$rm_new\

city_crime_train_mod$dis_new\


```{r, echo = FALSE, warning=FALSE, message=FALSE}

city_crime_train_mod$rm_new<-sin(city_crime_train$rm)

city_crime_train_mod$dis_new<-sin(city_crime_train$dis)

train_test_mod$rm_new<-sin(train_test$rm)

train_test_mod$dis_new<-sin(train_test$dis)

```

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(1,2))

#boxplot(city_crime_train_mod$rm_new,main="rm_new")
#boxplot(city_crime_train_mod$dis_new,main="dis_new")


```




##2.3 Tranformation for Variables

\

Following variables will need some transformation:

1. zn: proportion of residential land zoned for large lots (over 25000 square feet) (predictor variable) 
2. chas: a dummy var. for whether the suburb borders the Charles River (1) or not (0) (predictor variable) 
3. target: whether the crime rate is above the median crime rate (1) or not (0) (response variable) 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

city_crime_train_mod$zn_new<-city_crime_train_mod$zn

city_crime_train_mod$zn_new [city_crime_train_mod$zn_new > 0] <- 1
city_crime_train_mod$zn_new [city_crime_train_mod$zn_new== 0] <- 0
#city_crime_train_mod$zn_new<-factor(city_crime_train_mod$zn_new)



train_test_mod$zn_new<-train_test$zn

train_test_mod$zn_new [train_test_mod$zn_new > 0] <- 1
train_test_mod$zn_new [train_test_mod$zn_new== 0] <- 0
#city_crime_train_mod$zn_new<-factor(train_test_mod$zn_new)

#city_crime_train$chas <-factor(city_crime_train$chas)

#city_crime_train$target<-factor(city_crime_train$target)

#city_crime_train_mod$chas <-factor(city_crime_train_mod$chas)

#city_crime_train_mod$target<-factor(city_crime_train_mod$target)



```

##2.6 

\

Lets see how the new variables stack up against wins. 
\

```{r, echo = FALSE, warning=FALSE, message=FALSE}


#Correlation <- cor(city_crime_train$rm_new,(as.numeric(city_crime_train$target)))

#Correlation <- cor(city_crime_train$dis_new,(as.numeric(city_crime_train$target)))



```

\
All new variables seem to have a positive correlation with wins. However, some of them do not seem to have a strong correlation. Lets see how they perform while modeling.


\newpage


#3 Build Models




\newpage

Below is a summary table showing models and their respective variables. \

 
 
```{r, echo = FALSE, warning=FALSE, message=FALSE}

#modelvars <- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW1/ModelVars.csv")
#kable(modelvars)

```

\newpage 

##3.1.1 Model One by using all given variable
In this model, we will be using the original variables.  We will create model and we will highlight the variables that being recommended using the AIC value. \

First we will produce the summary model as per below: 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

model1 <- glm(target ~ ., data = city_crime_train, family = "binomial")

summary(model1)





#test<-(train_test$target==1)

#count(pre_train1>0.5==test)


```

Accuracy=0.9042553

###3.1.2 Model two- with backward step function with all given variables

```{r step for model 1}

stepmodel1<- step(model1, direction="backward")

pre_train1_step<-predict(stepmodel1,type="response",newdata=train_test)

table(pre_train1_step>0.5,train_test$target)



```
Accuracy=0.8723404



##3.1.3 Model three- model with transformed variables
In this model, we will be using the some transformed variables.  


First we will produce the summary model as per below: 

```{r, echo = FALSE, warning=FALSE, message=FALSE}

model2 <- glm(target ~ .-zn-rm-dis-tax-lstat-medv, data = city_crime_train_mod, family = "binomial")
summary(model2)

pre_train2<-predict(model2,type="response",newdata=train_test_mod)

test<-predict(model2,type="response")

table(pre_train2>0.5,train_test_mod$target)

#roc(test,city_crime_train$target)

# Accuracy


```

Accuracy=0.9042553


###3.1.4 Model with transformed variable and with with backward step function

```{r step for model 2}

stepmodel2<- step(model2, direction="backward")

pre_train2_step<-predict(stepmodel2,type="response",newdata=train_test_mod)

table(pre_train2_step>0.5,train_test_mod$target)

```

Accuracy= 0.893617

###3.1,5 Model three with Linear discrement analysis

```{r model with Linear Discrement analysis, eval=TRUE,echo=FALSE}


model3=lda(target~.,data=city_crime_train)


model3_predict_test<-data.frame(predict(model3,newdata=train_test))

model3_predict_test[1:5,]

table(model3_predict_test$class,train_test$target)

```

Accuracy=0.8297872


###3.1.6 Model with Linear discrement analysis with transformed data

```{r model with Linear Discrement analysis with modified data, eval=TRUE,echo=FALSE}


model3_mod=lda(target~.-zn-rm-dis-tax-lstat-medv,data=city_crime_train_mod)


model3_predict_test_mod<-data.frame(predict(model3,newdata=train_test_mod))



table(model3_predict_test_mod$class,train_test_mod$target)

```

Accuracy=0.7978723





#4 Model Selection 

In section we will further examine all six models.  We will apply a model selection strategy defined below to compare the models.
\


##4.1 Model selection strategy:

Following model selection strategy has been used for this assignment: 

(1) Compare accuracy of the models & confusion matrix
(2) Compare Precision,Sensitivity,Specificity,F1 score
(3) Compare AUC curve for the models

```{r model 1 performance,echo=FALSE}


#confusion matrix
pre_train1 <- predict(model1, newdata=train_test, type="response")
table(pre_train1>0.5,train_test$target)



#AUC
pred <- prediction(pre_train1, train_test$target)
perf <- performance(pred, measure = "tpr", x.measure = "fpr")
# I know, the following code is bizarre. Just go with it.
auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]

roc.data <- data.frame(fpr=unlist(perf@x.values),
                       tpr=unlist(perf@y.values),
                       model="GLM")
ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
    geom_ribbon(alpha=0.2) +
    geom_line(aes(y=tpr)) +
    ggtitle(paste0("ROC Curve w/ AUC=", auc))



```

