---
title: "Home Work Assignment - 01"
author:
- Critical Thinking Group 5
- Arindam Barman
- Mohamed Elmoudni
- Shazia Khan
- Kishore Prasad
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

\newpage


```{r, echo=FALSE , message=FALSE, warning=FALSE}
# changed lib path... 

# if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("MASS",character.only = TRUE)) (install.packages("MASS",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("knitr",character.only = TRUE)) (install.packages("knitr",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("xtable",character.only = TRUE)) (install.packages("xtable",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("psych",character.only = TRUE)) (install.packages("psych",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("stringr",character.only = TRUE)) (install.packages("stringr",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("car",character.only = TRUE)) (install.packages("car",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("faraway",character.only = TRUE)) (install.packages("faraway",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("aod",character.only = TRUE)) (install.packages("aod",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("ISLR",character.only = TRUE)) (install.packages("ISLR",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("AUC",character.only = TRUE)) (install.packages("AUC",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("ROCR",character.only = TRUE)) (install.packages("ROCR",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("leaps",character.only = TRUE)) (install.packages("leaps",repos = "http://cran.us.r-project.org", dep=TRUE))
# if (!require("pander",character.only = TRUE)) (install.packages("pander",repos = "http://cran.us.r-project.org", dep=TRUE))

library(ggplot2)
library(MASS)
library(knitr)
library(xtable)
# library(dplyr)
# library(psych)
library(stringr)
# library(car)
# library(faraway)
# library(aod)
# library(Rcpp)
# library(leaps)
# library(ISLR)
# library(AUC)
# library(ROCR)
# library(pander)

```

# Overview \

In this homework assignment, we will explore, analyze and model a data set containing information on approximately 12795 commercially available wines using 16 variables. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine. These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant. A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.


# Objective \

Our objective is to build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine. Using the training data set, we will build at least two different Poisson regression models, at least two different negative binomial regression models, and at least two multiple linear regression models, using different variables (or the same variables with different transformations).
 
To attain our objective, we will be following the below best practice steps and guidelines:

1 -Data Exploration \
2 -Data Preparation \
3 -Build Models \
4 -Select Models \


#1 Data Exploration Analysis

In section we will explore and gain some insights into the dataset by pursuing the below high level steps and inquiries: \
-Variable identification \
-Variable Relationships \
-Data summary analysis \
-Outliers and Missing Values Identification

##1.1	Variable identification


First we look the variables' datatypes and their roles.

```{r, echo=FALSE , message=FALSE, warning=FALSE}
#vartypes<- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW5/vartypes.csv")
#kable(vartypes)
```

From the Table 1 above, we see that that all variables are quantitative mainly of numeric and integer datatype. 
Also, we will ignore the INDEX variable as it is just a unique identifier for each row.  However, we will use the TARTGET variable as response variable and the remaining variables as predictors. 


##1.2 Variable Relationships \

Next let's display and examine the variable relationships as shown in table 2. \


```{r, echo=FALSE , message=FALSE, warning=FALSE}

winedata <- read.csv("https://raw.githubusercontent.com/kishkp/data621-ctg5/master/HW5//wine-training-data.csv")
#variables<- read.csv("C:/CUNY/Courses/IS621/Assignment602/Assignment05/vars.csv")
#kable(variables, caption = "Variable Description")

#pander::pander(variables, split.cells = c(20, 60, 40), split.table = Inf, justify = 'left', caption = "Variable Description")




```


At first glance, we can easily deduce that that the FreeSulfurDioxide (Sulfur Dioxide content of wine) can be derived from the TotalSulfurDioxide (Total Sulfur Dioxide of Wine). However, looking closer at the role of the sulfur dioxide $SO_2$, as it is used as a preservative because of its anti-oxidative and anti-microbial properties in wine and also as a cleaning agent for barrels and winery facilities, we realize that when a winemaker says his/her wine has 100 ppm (part per million) of $SO_2$, he/she is most probably referring to the total amount of $SO_2$ in his wine, and that means: \
total SO2 = free $SO_2$ + bound $SO_2$. \
free $SO_2$: molecular $SO_2$ + bisulfites + sulfites \
bound $SO_2$: sulfites attached to either sugars, acetaldehyde or phenolic compounds \

In this case the  free $SO_2$  portion (not associated with wine molecules) is effectively the buffer against microbes and oxidation... Hence without knowing the bound $SO_2$, we won't be able to derive  FreeSulfurDioxide from TotalSulfurDioxide.  

Also, looking breifly at the VolatileAcidity (Volatile Acid content of wine) and FixedAcidity (Fixed Acidity of Wine), we can easily deduce AcidIndex as the Acid index = Total acid (g/L) - pH. where Total acidity = Volatile Acid + Fixed Acidity. However, in our case the index is weighted average and we don't know the weighted average of either Volatile Acid or Fixed Acidity. Hence we will assume these variable do not have strict arithmetic relationships.

##1.3 Data summary analysis \

In this section, we will create summary data to better understand the initial relationship variables have with
our dependent variable using correlation, central tendency, and dispersion As shown in table 3.



```{r, echo=FALSE , message=FALSE, warning=FALSE}

library(dplyr)

winedata<- select(winedata, -(ï..INDEX))

ds_stats <- psych::describe(winedata, skew = FALSE, na.rm = TRUE)[c(3:6)]

ds_stats0<- ds_stats
ds_stats <- cbind(VARIABLE_NAME = rownames(ds_stats), ds_stats)
kable(ds_stats0, caption = "Data Summary")

```

\newpage
Below is the missing values and correlation table of the predictor variables to the response variables.   

```{r, echo=FALSE , message=FALSE, warning=FALSE}
Variable<- rownames(ds_stats)
fun <- function(x) sum(!complete.cases(x))
Missing <- sapply(winedata[Variable], FUN = fun) 
fun <- function(x, y) cor(y, x, use = "na.or.complete")
Correlation <- sapply(winedata[Variable], FUN = fun, y=winedata$TARGET) 
ds_stats2 <- data.frame(cbind( Missing, Correlation))
kable(ds_stats2, caption = "Missing Data and Data Correlation")
```

***Missing Values and Correlation Interpretation***

From tables 3 and 4 above, we observe the followings:

- Variable ResidualSugar has 616 and 0.0164913 correlation.  Given the low correlation we will try try some imputation techniques to handle the missing the values and replace missing values with their respective value. 
- variable Chlorides 638 -0.0382631 correlation. .  Given the low negative correlation we will try we would replace missing values with their respective value 
- Variable FreeSulfurDioxide 647 0.0438241. Given the low correlation we will impute the missing values with their respective value
- Variable TotalSulfurDioxide has 682 missing values with 0.0514784 correlation. Given the low correlation we will impute the missing values with their respective value.
- Variable Alcohol has 682 missing values with 0.0620616 correlation. Given the low correlation we will impute the missing values with their respective value. \


Please note that ResidualSugar, Chlorides, FreeSulfurDioxide, Alcohol, and TotalSulfurDioxide variables have similar number of missing values.  They are chemically related. However, we don't think they are arithmetically related. 

- In addition, variable pH has 395 missing values with negative correlation of  -0.0094448. Again we may just ignore these missing values especially that it has very low negative correlation to the target variable. 
- Variable Sulphates has much higher missing values of 1210 with low negative correlation of -0.0388496.  We will be imputing this values with their respective value 
- Now, variable STARS has the highest missing values of 3359 and highest correlation of 0.5587938. This is very important variable and it drives sales and consequently heavily impacts our response variable. 
We have to be careful in fixing the missing values as this variable STARS is rating score variable with 1 being the lowest and 4 the highest  


##1.4 Outliers Identification \

In this section we look at boxplots to determine the outliers in variables and decide on whether to act on the
outliers.
Lets do some univariate analysis. We will look at the Histogram and Boxplot for each variable to detect
outliers if any and treat it accordingly.

```{r, echo=FALSE , message=FALSE, warning=FALSE}

show_charts <- function(x, ...) {
    
    par(mfrow=c(2,3))
    
    xlabel <- unlist(str_split(deparse(substitute(x)), pattern = "\\$"))[2]
#    ylabel <- unlist(str_split(deparse(substitute(y)), pattern = "\\$"))[2]
    
    hist(x,main=xlabel)
    boxplot(x,main=xlabel)

    y<-log(x)
    boxplot(y,main='log transform')
    y<-sqrt(x)
    boxplot(y,main='sqrt transform')
    y<-sin(x)
    boxplot(y,main='sin transform')
    y<-(1/x)
    boxplot(y,main='Inverse transform')
}



```


```{r, echo=FALSE , message=FALSE, warning=FALSE}
show_charts(winedata$FixedAcidity)
show_charts(winedata$VolatileAcidity)
show_charts(winedata$ResidualSugar)
show_charts(winedata$Chlorides)
show_charts(winedata$FreeSulfurDioxide)
show_charts(winedata$TotalSulfurDioxide)
show_charts(winedata$Density)
show_charts(winedata$pH)
show_charts(winedata$Sulphates)
show_charts(winedata$Alcohol)
show_charts(winedata$LabelAppeal)
show_charts(winedata$AcidIndex)
show_charts(winedata$STARS)

```


\newpage

#2. Data Preparation 

Now that we have completed the preliminary analysis, we will be cleaning and consolidating data into one dataset for use in analysis and modeling. We will be puring the belwo steps as guidlines: \
- Outliers treatment \
- Missing values treatment \
- Data transformation \



##2.1 Outliers treatment

For outliers, we will create 2 sets of variables. 

The first set uses the capping method. In this method, we will replace all outliers that lie outside the 1.5 times of IQR limits. We will cap it by replacing those observations less than the lower limit with the value of 5th %ile and those that lie above the upper limit with the value of 95th %ile. 

Accordingly we create the following new variables while retaining the original variables. 


- FixedAcidity_CAP
- VolatileAcidity_CAP
- ResidualSugar_CAP
- Chlorides_CAP
- FreeSulfurDioxide_CAP
- TotalSulfurDioxide_CAP
- Density_CAP
- pH_CAP
- Sulphates_CAP
- Alcohol_CAP
- LabelAppeal_CAP
- AcidIndex_CAP
- STARS_CAP


```{r, echo = FALSE, warning=FALSE, message=FALSE}

treat_outliers <- function(x) {
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
 
return(x)
}

winedata$FixedAcidity_CAP <- treat_outliers(winedata$FixedAcidity)
winedata$VolatileAcidity_CAP <- treat_outliers(winedata$VolatileAcidity)
winedata$ResidualSugar_CAP <- treat_outliers(winedata$ResidualSugar)
winedata$Chlorides_CAP <- treat_outliers(winedata$Chlorides)
winedata$FreeSulfurDioxide_CAP <- treat_outliers(winedata$FreeSulfurDioxide)
winedata$TotalSulfurDioxide_CAP <- treat_outliers(winedata$TotalSulfurDioxide)
winedata$Density_CAP <- treat_outliers(winedata$Density)
winedata$pH_CAP <- treat_outliers(winedata$pH)
winedata$Sulphates_CAP <- treat_outliers(winedata$Sulphates)
winedata$Alcohol_CAP <- treat_outliers(winedata$Alcohol)
winedata$AcidIndex_CAP <- treat_outliers(winedata$AcidIndex)

```


\newpage


Lets see how the new variables look in boxplots.

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(2,3))

boxplot(winedata$FixedAcidity_CAP)
boxplot(winedata$VolatileAcidity_CAP)
boxplot(winedata$ResidualSugar_CAP)
boxplot(winedata$Chlorides_CAP)
boxplot(winedata$FreeSulfurDioxide_CAP)
boxplot(winedata$TotalSulfurDioxide_CAP)
boxplot(winedata$Density_CAP)
boxplot(winedata$pH_CAP)
boxplot(winedata$Sulphates_CAP)
boxplot(winedata$Alcohol_CAP)
boxplot(winedata$AcidIndex_CAP)

```


In the second set, we will use the sin transformation as identified in the data exploration and create the following variables:

- FixedAcidity_SIN
- VolatileAcidity_SIN
- ResidualSugar_SIN
- Chlorides_SIN
- FreeSulfurDioxide_SIN
- TotalSulfurDioxide_SIN
- Density_SIN
- pH_SIN
- Sulphates_SIN
- Alcohol_SIN
- AcidIndex_SIN

```{r, echo = FALSE, warning=FALSE, message=FALSE}

winedata$FixedAcidity_SIN <- sin(winedata$FixedAcidity)
winedata$VolatileAcidity_SIN <- sin(winedata$VolatileAcidity)
winedata$ResidualSugar_SIN <- sin(winedata$ResidualSugar)
winedata$Chlorides_SIN <- sin(winedata$Chlorides)
winedata$FreeSulfurDioxide_SIN <- sin(winedata$FreeSulfurDioxide)
winedata$TotalSulfurDioxide_SIN <- sin(winedata$TotalSulfurDioxide)
winedata$Density_SIN <- sin(winedata$Density)
winedata$pH_SIN <- sin(winedata$pH)
winedata$Sulphates_SIN <- sin(winedata$Sulphates)
winedata$Alcohol_SIN <- sin(winedata$Alcohol)
winedata$AcidIndex_SIN <- sin(winedata$AcidIndex)

```

```{r, echo = FALSE, warning=FALSE, message=FALSE}

par(mfrow=c(2,3))

boxplot(winedata$FixedAcidity_SIN)
boxplot(winedata$VolatileAcidity_SIN)
boxplot(winedata$ResidualSugar_SIN)
boxplot(winedata$Chlorides_SIN)
boxplot(winedata$FreeSulfurDioxide_SIN)
boxplot(winedata$TotalSulfurDioxide_SIN)
boxplot(winedata$Density_SIN)
boxplot(winedata$pH_SIN)
boxplot(winedata$Sulphates_SIN)
boxplot(winedata$Alcohol_SIN)
boxplot(winedata$AcidIndex_SIN)

```




##2.2 Missing values treatment

Next we impute missing values. Since we have handled outliers, we can go ahead and use the mean as impute values. As with outliers, we will go ahead and create new variables for the following:

- ResidualSugar
- Chlorides
- FreeSulfurDioxide
- TotalSulfurDioxide
- pH
- Sulphates
- Alcohol


```{r, echo = FALSE, warning=FALSE, message=FALSE}


winedata$ResidualSugar_IMP <- winedata$ResidualSugar 
winedata$Chlorides_IMP <- winedata$Chlorides
winedata$FreeSulfurDioxide_IMP <- winedata$FreeSulfurDioxide
winedata$TotalSulfurDioxide_IMP <- winedata$TotalSulfurDioxide
winedata$pH_IMP <- winedata$pH
winedata$Sulphates_IMP <- winedata$Sulphates
winedata$Alcohol_IMP <- winedata$Alcohol

winedata$ResidualSugar_IMP[is.na(winedata$ResidualSugar_IMP)] <- mean(winedata$ResidualSugar_IMP, na.rm = T) 
winedata$Chlorides_IMP[is.na(winedata$Chlorides_IMP)] <- mean(winedata$Chlorides_IMP, na.rm = T) 
winedata$FreeSulfurDioxide_IMP[is.na(winedata$FreeSulfurDioxide_IMP)] <- mean(winedata$FreeSulfurDioxide_IMP, na.rm = T) 
winedata$TotalSulfurDioxide_IMP[is.na(winedata$TotalSulfurDioxide_IMP)] <- mean(winedata$TotalSulfurDioxide_IMP, na.rm = T) 
winedata$pH_IMP[is.na(winedata$pH_IMP)] <- mean(winedata$pH_IMP, na.rm = T) 
winedata$Sulphates_IMP[is.na(winedata$Sulphates_IMP)] <- mean(winedata$Sulphates_IMP, na.rm = T) 
winedata$Alcohol_IMP[is.na(winedata$Alcohol_IMP)] <- mean(winedata$Alcohol_IMP, na.rm = T) 

```


Lets now create some additional variables that might help us in out analysis. 

##2.3 Missing Flags

We create flag variables to indicate whether some of the fields are missing any values. If the value is missing, we code it with 1 and if the value is present we code it with 0. The following are the variables that are created: 

- ResidualSugar
- Chlorides
- FreeSulfurDioxide
- TotalSulfurDioxide
- pH
- Sulphates
- Alcohol



```{r, echo = FALSE, warning=FALSE, message=FALSE}

winedata$ResidualSugar_MISS <- ifelse(is.na(winedata$ResidualSugar), 1, 0)
winedata$Chlorides_MISS <- ifelse(is.na(winedata$Chlorides), 1, 0)
winedata$FreeSulfurDioxide_MISS <- ifelse(is.na(winedata$FreeSulfurDioxide), 1, 0)
winedata$TotalSulfurDioxide_MISS <- ifelse(is.na(winedata$TotalSulfurDioxide), 1, 0)
winedata$pH_MISS <- ifelse(is.na(winedata$pH), 1, 0)
winedata$Sulphates_MISS <- ifelse(is.na(winedata$Sulphates), 1, 0)
winedata$Alcohol_MISS <- ifelse(is.na(winedata$Alcohol), 1, 0)

```
\
\


##2.4 Ratios

\

Next we create some additional variables, that we think may be useful with the prediction. Here we create the following ratios: \

Hits_R  = TEAM_BATTING_H/TEAM_PITCHING_H \
Walks_R = TEAM_BATTING_BB/TEAM_PITCHING_BB \
HomeRuns_R = TEAM_BATTING_HR/TEAM_PITCHING_HR \
Strikeout_R = TEAM_BATTING_SO/TEAM_PITCHING_SO \



```{r, echo = FALSE, warning=FALSE, message=FALSE}

# moneyball2$Hits_R <- moneyball2$TEAM_BATTING_H/moneyball2$TEAM_PITCHING_H
# moneyball2$Walks_R <- moneyball2$TEAM_BATTING_BB/moneyball2$TEAM_PITCHING_BB
# moneyball2$HomeRuns_R <- moneyball2$TEAM_BATTING_HR/moneyball2$TEAM_PITCHING_HR
# moneyball2$Strikeout_R <- moneyball2$TEAM_BATTING_SO/moneyball2$TEAM_PITCHING_SO


```

\
\


##2.5 Calculated Variables

\

Finally, we will also create calculated variables as below:

1. TEAM_BATTING_EB (Extra Base Hits) = 2B + 3B + HR
2. TEAM_BATTING_1B (Singles by batters) = TEAM_BATTING_H - TEAM_BATTING_EB

```{r, echo = FALSE, warning=FALSE, message=FALSE}

# moneyball2$TEAM_BATTING_EB <- moneyball2$TEAM_BATTING_2B + moneyball2$TEAM_BATTING_3B + moneyball2$TEAM_BATTING_HR
# 
# moneyball2$TEAM_BATTING_1B <- moneyball2$TEAM_BATTING_H - moneyball2$TEAM_BATTING_EB

```

##2.6 Correlation for new variables

\

Lets see how the new variables stack up against wins. 
\

```{r, echo = FALSE, warning=FALSE, message=FALSE}

fun <- function(x, y) cor(y, x, use = "na.or.complete")
Correlation <- sapply(winedata[, 16:ncol(winedata)], FUN = fun, y=winedata$TARGET) 
#kable(Correlation, caption = "New variables Correlation ")
Correlation
```

\
All new variables seem to have a positive correlation with wins. However, some of them do not seem to have a strong correlation. Lets see how they perform while modeling.
